---
title: "Stock Market Analysis"
author: "Sagar Chettiyar and Bhakti Sangoi"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Required Packages
##Data Manipulation Packages
```{r}
library("tidyverse")
library("dplyr")
library("lubridate")
library("splitstackshape")
```

##Graphic Packages
```{r}
library("ggthemes")
library("highcharter")
library("ggplot2")
library("plotly")
```

##SQLlite Package
```{r}
library("RSQLite")
```

##Finance Package
```{r}
library("quantmod")
```

##Forecasting Packages
```{r}
library("tseries")
library("StanHeaders")
library("rstan")
library("prophet")
library("forecast")
library("TSPred")
```

#DATA COLLECTION
```{r}
Company_Info<-read_csv("Comapny_Information.csv")
Annual_Balance_Info<-read_csv("Annual_Balance_Sheet.csv")
Daily_Stock_Info<-read_csv("Daily_Stock_Prices.csv")
```

#DATA CLEANING

##For Company Info
```{r}
#Splitting the address into City and State
Company_Info<-cSplit(Company_Info, "Address of Headquarters", ",", fixed = FALSE)

#Dropping unwanted columns
Company_Info$`SEC filings`<-NULL
Company_Info$`Address of Headquarters_3`<-NULL

#Renaming the dataset for better understanding
names(Company_Info)<-c("Ticker_Symbol",
                       "Company_Name",
                       "GICS_Sector",
                       "GICS_Major_Industry",
                       "Date_Added",
                       "CIK",
                       "City",
                       "State")

#Filling Missing values
#Replacing the NA's in date column with a default date:"1999-01-01"
Company_Info$Date_Added[is.na(Company_Info$Date_Added)] <- "1999-01-01"

#Replacing IT with Information Technology in GICS Sector
Company_Info$GICS_Sector[Company_Info$GICS_Sector=="IT"] <- "Information Technology"
```

##For Daily_Stock_Info
```{r}
#Renaming the dataset for better understanding
names(Daily_Stock_Info)<-c("Date",
                           "Ticker_Symbol",
                           "Open_Price",
                           "Close_Price",
                           "Lowest_Price",
                           "Highest_Price",
                           "Volume")

#Subsetting date column to keep only dates
Daily_Stock_Info_Table<-Daily_Stock_Info #Backing up the dataframe for SQL Use
#Factorizing Dates
Daily_Stock_Info$Date<-as.Date(Daily_Stock_Info$Date, "%m/%d/%Y") #Reformatting Date
Daily_Stock_Info$Day<-factor(day(as.POSIXlt(Daily_Stock_Info$Date, format="%m/%d/%Y"))) #Adding a Day column representing the no. of day of the month
Daily_Stock_Info$Month<-factor(month(as.POSIXlt(Daily_Stock_Info$Date, format="%m/%d/%Y"))) #Adding a Month column representing the Month in Numbers
Daily_Stock_Info$Year<-factor(year(as.POSIXlt(Daily_Stock_Info$Date, format="%m/%d/%Y"))) #Seperating the Year from the date field
Daily_Stock_Info$Weekday<-factor(wday(as.POSIXlt(Daily_Stock_Info$Date, format="%m/%d/%Y"))) #Adding a weekday column representing the day of the week

#Creating new dataframes for future analysis
Volume_Year<-Daily_Stock_Info %>% 
  filter(Year %in% c("2012","2013","2014","2015","2016")) %>% 
  group_by(Year) %>% 
  summarise(Total = n())
#This dataframe contains volume of stocks traded by  Year

Volume_Month<-Daily_Stock_Info %>% 
          group_by(Month) %>% 
          summarise(Total = n())
Month<-c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec") #Assigning month names instead of month numbers
Volume_Month$Month<-Month[Volume_Month$Month]
#This dataframe contains volume of stocks traded by Month

Volume_YearMonth<-Daily_Stock_Info %>% 
  group_by(Year, Month) %>% 
  summarise(Total = n())
#This dataframe contains volume of stocks traded by Month and Year

Volume_Weekday<-Daily_Stock_Info %>% 
          group_by(Weekday) %>% 
          summarise(Total = n())
Weekday<-c("Monday","Tuesday","Wednesday","Thursday","Friday")  #Assigning Weekday names instead of month numbers
#Note: Saturday and Sunday are not included because the stock market does not remain functional on these days
Volume_Weekday$Weekday<-Weekday[Volume_Weekday$Weekday]
#This dataframe contains volume of stocks traded by weekdays

Volume_Day<-Daily_Stock_Info %>% 
          group_by(Day) %>% 
          summarise(Total = n())
#This dataframe contains volume of stocks traded by each day of the month
```

##For Annual_Balance_Info
```{r}
#Renaming the dataset for better understanding
names(Annual_Balance_Info)<-c("Financial_Year",
                       "Ticker_Symbol",
                       "Cash_And_Cash_Equivalents",
                       "Total_Current_Assets",
                       "Total_Current_Liabilities",
                       "Short_Term_Investments",
                       "Total_Revenue",
                       "Treasury_Stock",
                       "Retained_Earnings",
                       "Operation_Margin",
                       "Quick_Ratio")
```

#DATA STORAGE
```{r}
#Creating Database Connection
db_conn <- dbConnect(SQLite(), dbname="Stock_Market.sqlite")
#dbDisconnect(db_conn)

#Table Creation
#For Company_Info
dbGetQuery(db_conn, #Database Connector name
           "create table Company_Info
           (
           Ticker_Symbol Text Primary Key,
           Company_Name Text,
           GICS_Sector Text,
           GICS_Major_Industry Text,
           Date_Added Date,
           CIK Text,
           City Text,
           State Text
           )") #Create Table Script
dbWriteTable(conn = db_conn, #Database Connector name
             name = "Company_Info", #Table Name
             value = Company_Info, #Load data from the newly created dataframe
             append=TRUE,row.names = FALSE,header = TRUE)#Since the CSV contains headers in the columns
dbListFields(db_conn, "Company_Info") #Listing the field of the table


#For Daily_Stock_Info
dbGetQuery(db_conn, #Database Connector name
           "create table Daily_Stock_Info
           (
           Daily_Stock_ID Integer Primary Key Autoincrement,
           Date Date,
           Ticker_Symbol Text,
           Open_Price Numeric,
           Close_Price Numeric,
           Lowest_Price Numeric,
           Highest_Price Numeric,
           Volume Real
           )") #Create Table Script
dbWriteTable(conn = db_conn, #Database Connector name
             name = "Daily_Stock_Info", #Table Name
             value = Daily_Stock_Info_Table, #Load data from the newly created dataframe
             append=TRUE,row.names = FALSE,header = TRUE)#Since the CSV contains headers in the columns
dbListFields(db_conn, "Daily_Stock_Info") #Listing the field of the table

#For Annual_Balance_Info
dbGetQuery(db_conn, #Database Connector name
           "create table Annual_Balance_Info
           (
           Annual_Stock_ID Integer Primary Key Autoincrement,
           Financial_Year Integer,
           Ticker_Symbol Text,
           Cash_And_Cash_Equivalents Numeric,
           Total_Current_Assets Numeric,
           Total_Current_Liabilities Numeric,
           Short_Term_Investments Numeric,
           Total_Revenue Numeric,
           Treasury_Stock Numeric,
           Retained_Earnings Numeric,
           Operation_Margin Numeric,
           Quick_Ratio Numeric
           )") #Create Table Script
dbWriteTable(conn = db_conn, #Database Connector name
             name = "Annual_Balance_Info", #Table Name
             value = Annual_Balance_Info, #Load data from the newly created dataframe
             append=TRUE,row.names = FALSE,header = TRUE)#Since the CSV contains headers in the columns
dbListFields(db_conn, "Annual_Balance_Info") #Listing the field of the table
```

#DATA RETRIEVAL
#1) Display the companies having current ratio greater than one
```{r}
query<-"select a.Ticker_Symbol,c.Company_Name, (a.Total_Current_Assets/a.Total_Current_Liabilities) as Current_Ratio
        from Annual_Balance_Info a
        join Company_Info c on a.Ticker_Symbol=c.Ticker_Symbol
        group by a.Ticker_Symbol,c.Company_Name
        having (a.Total_Current_Assets/a.Total_Current_Liabilities)>1
        order by (a.Total_Current_Assets/a.Total_Current_Liabilities) desc
        limit 10"
dbGetQuery(db_conn,query) #Calling dbquery to run the query on db_conn and show output
```

#2)Which state has the highest number of headquarters?
```{r}
query<-"select State,count(Ticker_Symbol) as No_Of_Headquarters
        from Company_Info
        group by State
        order by count(Ticker_Symbol) desc
        limit 5"
dbGetQuery(db_conn,query) #Calling dbquery to run the query on db_conn and show output
```

#3)Display all the GICS Sector along with their Total revenue
```{r}
query<-"select c.GICS_Sector, sum(a.Total_Revenue) as Total_Revenue
        from Annual_Balance_Info a
        join Company_Info c on a.Ticker_Symbol=c.Ticker_Symbol
        group by c.GICS_Sector
        order by sum(a.Total_Revenue) desc"
dbGetQuery(db_conn,query) #Calling dbquery to run the query on db_conn and show output
```

#4)When was the highest volume of stocks traded? For which company? How much?
```{r}
query<-"select d.Date, c.Company_Name, max(d.Volume) as Max_Volume
        from Daily_Stock_Info d
        join Company_Info c on d.Ticker_Symbol=c.Ticker_Symbol"
dbGetQuery(db_conn,query) #Calling dbquery to run the query on db_conn and show output
```

#5)Which Industry has the highest number of stocks traded?
```{r}
query<-"select c.GICS_Sector, count(d.Volume) as No_Of_Stocks_Traded
        from Daily_Stock_Info d
        join Company_Info c on d.Ticker_Symbol=c.Ticker_Symbol
        group by c.GICS_Sector
        order by count(d.Volume) desc"
dbGetQuery(db_conn,query) #Calling dbquery to run the query on db_conn and show output
```

#6)Display the Retained earnings for Industries under Telecommunications Sector?
```{r}
query<-"select c.GICS_Major_Industry, sum(a.Retained_Earnings) as Total_Retained_Earnings
        from Annual_Balance_Info a
        join Company_Info c on a.Ticker_Symbol=c.Ticker_Symbol
        group by c.GICS_Major_Industry
        having c.GICS_Sector='Consumer Discretionary'
        order by sum(a.Retained_Earnings) desc"
dbGetQuery(db_conn,query) #Calling dbquery to run the query on db_conn and show output
```

#DATA ANALYSIS
#1)What are the Total Current Liabilities and Assets of American Airlines Group?
```{r}
#Calculations
Annual_Info_CR_Ratio<-Annual_Balance_Info %>% 
                     select(Financial_Year,Ticker_Symbol,Total_Current_Liabilities,Total_Current_Assets)
Annual_Info_CR_Ratio$Current_Ratio<-Annual_Info_CR_Ratio$Total_Current_Assets/Annual_Info_CR_Ratio$Total_Current_Liabilities #Calculating the Current Ratio for all the stocks
Annual_Profit<-filter(Annual_Info_CR_Ratio, Current_Ratio>=1) #This dataframe contains all the companies that have higher assets than liabilities
Annual_Loss<-filter(Annual_Info_CR_Ratio, Current_Ratio<1) #This dataframe contains all the companies that have more liabilities than assets

Annual_Balance_Info$Cash_Ratio<-Annual_Balance_Info$Cash_And_Cash_Equivalents/Annual_Balance_Info$Total_Current_Liabilities #Calculating the Cash Ratio for all the stocks

AAL_Info<-Annual_Info_CR_Ratio %>% 
    select(Financial_Year,Ticker_Symbol,Total_Current_Liabilities,Total_Current_Assets,Current_Ratio) %>% 
    filter(Ticker_Symbol=='AAL') #Filtering for American Airlines Group
AAL_Info

highchart() %>%
  hc_xAxis(categories = AAL_Info$Financial_Year) %>%
  hc_title(text = "Total_Current_Liabilites vs Total_Currents_Assets for American Airlines Group from 2012 to 2015") %>% #Title of the graph
  hc_add_series(
    name='Total_Current_Liabilities', #Legend
    color='red',#Color of the line graph
    data=AAL_Info$Total_Current_Liabilities #Column to Display
  ) %>%
  hc_add_series(
    name='Total_Current_Assets', #Legend
    color='green', #Color of the line graph
    data=AAL_Info$Total_Current_Assets #Column to Display
  )
```

#2)Plot a graph of how much a company make's on each dollar of sales [Operation Margin].
```{r}
Opt_Margin<-Annual_Balance_Info%>%
  select(Ticker_Symbol,Operation_Margin)%>%
  arrange(desc(Operation_Margin))%>% #Arranging in descending order 
  head(n=10)%>% #Selecting top 10 Highest Operation Margin
  distinct() #Selecting only distinct values as the Ticker_Symbl may repeat because of multiple financial year

hchart(Opt_Margin,'column', hcaes(x = Ticker_Symbol, y = Operation_Margin, color = Operation_Margin)) %>%  #Plot type Column
  hc_add_theme(hc_theme_flat()) %>%  #Using flat theme for appealing visualization
  hc_title(text="Top 10 stocks with highest Operation Margin") #Graph Title
```

#3)Display the top 5 states having higher number of headquarters.
```{r}
State_HQ<-Company_Info%>% 
          select(State)%>%
          group_by(State)%>%
          summarise(No_of_HQ = n())%>% #Counting the number of HeadQuarters
          arrange(desc(No_of_HQ)) #Arranging in descending order [Max No. of HQ on top]

hchart(State_HQ[1:5,],'pie', hcaes(x = State, y = No_of_HQ, color = No_of_HQ)) %>%  #Plot type pie and selecting only top 5 records
  hc_add_theme(hc_theme_google()) %>%  #Theming for better visualization
  hc_title(text="Top 5 states having higher number of headquarters") #Graph Title
```

#4)Volume of Stocks traded by Year and Month[2010-2016]
```{r}
Volume_YearMonthPlot<-ggplot(Volume_YearMonth, aes(Year, Month, fill = Total)) + #Using the fill option as the Total for the blocks
         geom_tile(size = 1, color = "white") + #Setting the size and color of tile
         theme_classic() + #Theme
         geom_text(aes(label=Total), color='white') + #Tile labels are the Total Value
         ggtitle("Volume of Stocks traded by Year and Month[2010-2016]")  #Graph Title
plot(Volume_YearMonthPlot) 
```

```{r}
#5)Volume of Stocks traded by Month
hchart(Volume_Month,'line', hcaes(x = Month, y = Total, color = Total)) %>% #Line plot
  hc_title(text = "Volume of Stocks traded by month of the year")  #Title
```

#6)Volume of Stocks traded by each day of the month
```{r}
hchart(Volume_Day,'column', hcaes(x = Day, y = Total, color = Total)) %>%  #Column Plot
  hc_title(text = "Volume of Stocks traded by each Day of the month") #Title
```

#7)Volume of Stocks traded by Weekday
```{r}
hchart(Volume_Weekday,'line', hcaes(x = Weekday, y = Total, color = Total)) %>%  #Line Plot
  hc_title(text = "Volume of Stocks traded by Day of the Week") #Title
```

#8)Display the Total Revenue for Energy,Financials,Information Technology,Health Care,Consumer Staples and Industrials from 2012 to 2015
```{r}
query<-"select a.Financial_Year,c.GICS_Sector, sum(a.Total_Revenue) as Total_Revenue
        from Annual_Balance_Info a
join Company_Info c on a.Ticker_Symbol=c.Ticker_Symbol
group by a.Financial_Year,c.GICS_Sector
order by a.Financial_Year,c.GICS_Sector"

GICS<-dbGetQuery(db_conn,query) #Calling dbquery to run the query on db_conn and show output and storing it for future reference
GICS<-na.omit(GICS) #Removing Nulls
GICS<-GICS%>% filter(Financial_Year %in% c("2012","2013","2014","2015")) #Filtering out for the period 2012-2015
GICS<-GICS%>% filter(GICS_Sector %in% c("Energy",
                                  "Financials",
                                  "Information Technology",
                                  "Health Care",
                                  "Consumer Staples",
                                  "Industrials"))
#Choosing only Top GICS sector for clear visualization

hchart(GICS,'column', hcaes(x = Financial_Year, y = Total_Revenue, group = GICS_Sector)) %>% 
  hc_add_theme(hc_theme_google()) %>% 
  hc_plotOptions(column = list(stacking = 'normal')) %>% 
  hc_legend(align = 'right', float = T)%>% 
  hc_title(text = "Total Revenue as per GICS Sector for each Financial Year") 
```

#9) Which companies have more Short Term Investments Due compared to Cash and Cash Equivalents
```{r}
query<-"select Ticker_Symbol,Short_Term_Investments,Cash_And_Cash_Equivalents, Short_Term_Investments - Cash_And_Cash_Equivalents
        from Annual_Balance_Info
        group by Ticker_Symbol
        having Short_Term_Investments > Cash_And_Cash_Equivalents and Short_Term_Investments > 0 and Cash_And_Cash_Equivalents > 0
        order by (Short_Term_Investments - Cash_And_Cash_Equivalents) desc
        limit 10"
Loss_Stock<-dbGetQuery(db_conn,query) #Calling dbquery to run the query on db_conn and show output

plot_ly(Loss_Stock, x = Loss_Stock$Ticker_Symbol, y = Loss_Stock$Short_Term_Investments, type = 'bar', name = 'Short Term Investments Due') %>%
  add_trace(y = Loss_Stock$Cash_And_Cash_Equivalents, name = 'Cash and Cash Equivalents') %>% #Adding another group column for the graph
  layout(yaxis = list(title = 'Amount in Dollars'), barmode = 'group',title='Comapnies with higher Short Term Investments Due compared to Cash and Cash Equivalents ') #Title for graph
```

#TIME SERIES ANALYSIS
```{r}
#Using quantmod package gathering live data for the following companies
Google<-getSymbols(Symbols="GOOG", src = "yahoo", auto.assign = FALSE, verbose = TRUE) #Pulls data from Yahoo Finance for Google Stocks
Amazon<-getSymbols(Symbols="AMZN", src = "yahoo", auto.assign = FALSE, verbose = TRUE) #Pulls data from Yahoo Finance for Amazon Stocks
Microsoft<-getSymbols(Symbols="MSFT", src = "yahoo", auto.assign = FALSE, verbose = TRUE) #Pulls data from Yahoo Finance for Microsoft Stocks
Facebook<-getSymbols(Symbols="FB", src = "yahoo", auto.assign = FALSE, verbose = TRUE) #Pulls data from Yahoo Finance for Facebook Stocks

highchart(type="stock") %>% 
  hc_title(text="Stock Analyses of the major Tech Giants") %>%   
  hc_add_series(Google) %>% 
  hc_add_series(Amazon) %>% 
  hc_add_series(Microsoft) %>% 
  hc_add_series(Facebook,type="ohlc")
```

#FORECASTING
##Prophet(Forecasting the Volume of Stocks)
```{r}
#Step1:Collect Data
Volume_Date<-Daily_Stock_Info %>% 
             group_by(Date) %>%  #Grouping data by date
             summarise(Total = n()) #Volume of stocks

#Step2:Calculate the logarithmic parameter
names(Volume_Date)<-c("ds","y")
Volume_Date<-Volume_Date %>%
             mutate(y = log(y))
Volume_Prophet<-prophet(Volume_Date)

#Step3:Forecasting
Future<-make_future_dataframe(Volume_Prophet,periods = 730) #Period = 730 because we are forecastig it for two years
Forecast_Prophet<-predict(Volume_Prophet,Future)
plot(Volume_Prophet,Forecast_Prophet,ylab="Volume of stocks",xlab="Year") #Plotting a graph for the Volume of stocks with Year
prophet_plot_components(Volume_Prophet,Forecast_Prophet) #Plotting other forecasting components such as trends, weekly and yearly
```

##ARIMA Model(Forecasting the High Price of Stocks)
```{r}
#Step1:Prepare Data
Volume_Date_ARIMA<-Daily_Stock_Info %>% 
  group_by(Date) %>%  #Grouping data by date
  summarise(Total = n()) #Volume of stocks

#Step2:Creating Time Series
Volume_Date_ARIMA = ts(na.omit(Volume_Date_ARIMA$Total), start=c(2016,1), end=c(2018,12),frequency=24) #Preparing time series
plot(Volume_Date_ARIMA)#Plot of the High Price of the stocks

#Step3:Decomposing the Data
Decomp = stl(Volume_Date_ARIMA, s.window="periodic") #STL is a flexible function for decomposing and forecasting the series. 
Deseasonal<-seasadj(Decomp) #Returns seasonally adjusted data constructed by removing the seasonal component.
plot(Decomp)

#Step4:Stationary
adf.test(Volume_Date_ARIMA,alternative="stationary") #ADF procedure tests whether the change in Y can be explained by lagged value and a linear trend.

#Step5:Autocorrelations and Choosing Model Order
Differenced_Series = diff(Deseasonal,differences = 1)
plot(Differenced_Series)
adf.test(Differenced_Series,alternative="stationary")

acf(Differenced_Series,main='ACF for Differenced Series')
pacf(Differenced_Series,main='PACF for Differenced Series')

#Step6:Fitting ARIMA Model
Arima=auto.arima(Volume_Date_ARIMA,trace = TRUE,test="kpss",ic="aic") #Fits the model using approximation and no-approximation and picks the Best Model
plot.ts(Arima$residuals) #Plots the Residual component of the ARIMA

acf(Arima$residuals,lag.max=54)
pacf(Arima$residuals,lag.max=54)

Arima_Forecast = forecast(Arima, h=24) #Forecasts the model for 24 months
Arima_Forecast
plot(Arima_Forecast, xlab="Time") #Plots the forecast graph for the best model selected above by Auto_ARIMA

```